\documentclass{article}

% NeurIPS 2026 Data & Benchmark Workshop
% TODO: Update to neurips_2026 style when available
\usepackage[final]{neurips_2024}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,fit,calc}
\usepackage{listings}
\usepackage{float}

% Code listing style
\lstset{
  language=Python,
  basicstyle=\ttfamily\footnotesize,
  keywordstyle=\color{blue!70!black},
  commentstyle=\color{green!50!black},
  stringstyle=\color{red!60!black},
  showstringspaces=false,
  breaklines=true,
  frame=single,
  framesep=3pt,
  xleftmargin=6pt,
  xrightmargin=6pt,
  aboveskip=6pt,
  belowskip=4pt,
}

% Custom commands
\newcommand{\heron}{\textsc{Heron}}
\newcommand{\ie}{\textit{i.e.}}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\etal}{\textit{et al.}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\newcommand{\placeholder}[1]{\textcolor{blue}{[PLACEHOLDER: #1]}}
\newcommand{\exampleval}[1]{\textcolor{teal}{#1}\textsuperscript{\textcolor{teal}{$\dagger$}}}
\newcommand{\cmark}{\checkmark}
\newcommand{\xmark}{$\times$}

\title{\heron{}: Agent-Centric Multi-Agent Simulation\\for Event-Driven Hierarchical Coordination}

\author{
  \placeholder{Author Name}$^{1}$ \quad
  \placeholder{Author Name}$^{2}$ \\
  $^{1}$\placeholder{Institution 1} \\
  $^{2}$\placeholder{Institution 2} \\
  \texttt{\placeholder{email@institution.edu}}
}

\begin{document}

\maketitle

\begin{abstract}
Key MARL research questions---Does asynchronous execution degrade cooperation? How much information must agents share for stable coordination?---remain understudied because answering them requires independently varying execution model (when agents act), information structure (what they observe), and coordination mechanism (how they interact). Standard frameworks conflate all three inside a single \texttt{step()} loop; changing any one means modifying the same code that handles physics, introducing confounds that standard ablations cannot isolate.

We introduce \heron{}, a benchmark framework that disentangles these three dimensions: \emph{agents own the coordination; the environment owns only the physics}. \heron{} provides (1)~a three-level agent hierarchy with composable coordination protocols, (2)~dual-mode execution---synchronous \texttt{step()} for CTDE training and event-driven \texttt{run\_event\_driven()} for deployment-realistic evaluation, and (3)~a four-level visibility model (\texttt{public}/\texttt{owner}/\texttt{upper\_level}/\texttt{system}) that we use to show upper-level visibility achieves near-optimal performance (+3.7\% cost) while respecting realistic information boundaries. We further show that policies trained under synchronous assumptions suffer \placeholder{X}\% degradation under communication delays calibrated to power-system SCADA timescales (seconds to minutes), and that these ordering effects hold across MAPPO, IPPO, and QMIX. We demonstrate on power systems (6 IEEE/CIGRE networks) and traffic signal control; both domains reuse identical base abstractions without any domain-specific coordination code.
\end{abstract}

%==============================================================================
\section{Introduction}
\label{sec:intro}
%==============================================================================

MARL algorithms are typically developed in environments that share a common execution model: a centralized loop calls \texttt{env.step(actions)}, delivers observations derived from a global state snapshot, and collects actions from all agents simultaneously. This \emph{synchronous, environment-centric} paradigm---inherited from single-agent RL and extended by PettingZoo~\cite{terry2021pettingzoo}, EPyMARL~\cite{papoudakis2021benchmarking}, and MARLlib~\cite{hu2023marllib}---has been productive for algorithm benchmarking. However, it encodes three assumptions that collapse in cyber-physical systems (CPS) such as smart grids, traffic networks, and robotic fleets.

\paragraph{Three assumptions that break.}
\begin{enumerate}[nosep,leftmargin=*]
    \item \textbf{Synchronized execution.} All agents observe and act at the same instant. In reality, agents operate at heterogeneous rates---a field sensor polls every second, a zone coordinator decides every minute. Observations are stale by the time they arrive; actions take effect after physical delays.
    \item \textbf{Global information access.} Observations are derived from a single global state snapshot, differing across agents only by what the environment \emph{chooses} to mask. In reality, information is partitioned by organizational boundaries, sensor coverage, and communication topology.
    \item \textbf{Implicit coordination.} Agents coordinate through shared reward signals and centralized critics. In reality, coordination requires explicit mechanisms---setpoint commands flowing down a hierarchy, price signals propagating through a market, consensus protocols running among peers---each with its own latency and failure modes.
\end{enumerate}

Training a policy under MDP assumptions (shared observations, simultaneous actions) and deploying it under Dec-POMDP conditions (asymmetrically stale observations, delayed effects) exposes a structural mismatch: the optimization landscape during training does not match the decision problem at deployment.

\paragraph{Why existing frameworks don't solve this.}
Standard MARL frameworks standardize the \emph{environment-algorithm interface} and do this well. But none standardize \emph{what happens inside the environment}---how agents access state, how they communicate, or how execution timing affects observations. When a researcher needs to restrict a field agent's observation, compare centralized versus decentralized coordination, or measure timing-induced degradation, they must build these mechanisms \emph{inside the environment class}---modifying the same code that handles physics. Each variation requires a new environment implementation. Subtle differences in physics handling between variants introduce confounds that do not appear in any configuration table. You think you are comparing coordination structures, but you are comparing two partially divergent simulators.

\paragraph{The agent-centric alternative.}
\heron{} inverts the locus of control. Instead of the environment owning all agent logic, \heron{} places coordination in \textbf{agents and their protocols}:
\begin{itemize}[nosep]
    \item \textbf{The environment owns the physics}---three bridge methods connect \heron{}'s state representation to any external simulator. Everything else is stable across experiments.
    \item \textbf{Agents own the coordination}---agents are explicitly hierarchical (\texttt{FieldAgent} $\rightarrow$ \texttt{CoordinatorAgent} $\rightarrow$ \texttt{SystemAgent}), and coordination is defined by pluggable \texttt{Protocol} objects. Swapping from centralized dispatch to peer consensus is a constructor argument change, not an environment rewrite.
    \item \textbf{The scheduler owns the timing}---a heap-based \texttt{EventScheduler} advances simulation time. The same agent hierarchy runs in both synchronous and event-driven modes, with per-agent \texttt{TickConfig} specifying intervals, delays, and jitter distributions.
\end{itemize}

\paragraph{Contributions.}
\heron{} is a benchmark framework with contributions around a single theme---\emph{making execution model, information structure, and coordination mechanisms first-class design variables}:
\begin{enumerate}[leftmargin=*, itemsep=2pt]
    \item \textbf{Agent-centric architecture} (\S\ref{sec:hierarchy}): A three-level agent hierarchy where physics, coordination, and timing are cleanly separated.
    \item \textbf{Dual-mode execution} (\S\ref{sec:concepts}): Synchronous \texttt{env.step()} for CTDE training and event-driven \texttt{env.run\_event\_driven()} via a heap-based \texttt{EventScheduler} with 7 event types---enabling direct measurement of the synchronous-to-event-driven gap.
    \item \textbf{ProxyAgent for state mediation} (\S\ref{sec:proxy}): A centralized gatekeeper enforcing four-level visibility constraints.
    \item \textbf{Composable protocol system} (\S\ref{sec:protocols}): Orthogonal \texttt{CommunicationProtocol} + \texttt{ActionProtocol} composition.
    \item \textbf{Power systems case study} (\S\ref{sec:power}): 20 FeatureProviders, 6 agent types, 6 IEEE/CIGRE test networks.
\end{enumerate}

%==============================================================================
\section{Concepts: Event-Driven Execution as a Full Episode Lifecycle}
\label{sec:concepts}
%==============================================================================

\subsection{Step-Based vs. Event-Driven Execution}

In conventional MARL, an episode is a sequence of \emph{synchronized global steps}: at each step, all agents see the same time snapshot, decide simultaneously, and the environment advances one tick. Rewards $r_t$ are attached to individual steps. The decision process is a (PO)MDP. This is correct for CTDE training.

In \emph{event-driven} execution, an episode is a continuous simulation over $[0, t_\text{end}]$ where agents act asynchronously on independent clocks. A heap-based scheduler processes events chronologically:
\begin{itemize}[nosep]
    \item On \texttt{AGENT\_TICK}: the agent observes (potentially stale) state, decides, and schedules an \texttt{ACTION\_EFFECT} with delay.
    \item On \texttt{ACTION\_EFFECT}: the action takes physical effect, updating the state cache.
    \item On \texttt{MESSAGE\_DELIVERY}: a message arrives at the recipient after communication delay.
\end{itemize}

Table~\ref{tab:exec-comparison} summarizes the differences.

\begin{table}[t]
\centering
\caption{Step-based vs.\ event-driven execution semantics.}
\label{tab:exec-comparison}
\small
\begin{tabular}{@{}lp{4.8cm}p{5.2cm}@{}}
\toprule
\textbf{Dimension} & \textbf{Step-based} & \textbf{Event-driven} \\
\midrule
Observation timing & All agents see the same snapshot at time $t$ & Each agent sees whatever was last written to the state cache---different historical states \\
Decision process & Shared-observation (PO)MDP & Dec-POMDP with asymmetrically stale observations \\
Action effect & Immediate, simultaneous & Delayed by \texttt{act\_delay} \\
Communication & Implicit via shared state & Explicit messages with \texttt{msg\_delay} \\
Agent cadence & All tick together & Independent \texttt{tick\_interval} with optional jitter \\
Reward semantics & Per-step $r_t$ & Hierarchical cascade; aggregated over episode \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Why Both Modes in One Framework}

The core scientific question \heron{} enables is: \emph{how much does policy performance degrade when we move from the training-time MDP to the deployment-time Dec-POMDP?} This requires both modes over the \emph{same} agent definitions: train with \texttt{env.step()} and evaluate with \texttt{env.run\_event\_driven()}. The reward difference quantifies the \emph{synchronous-to-event-driven gap}, which depends on the agent hierarchy, timing configuration, and information structure---all independently tunable.

\subsection{Rewards in Event-Driven Mode}

In step-based mode, rewards are computed per-step and returned immediately. In event-driven mode, rewards follow a \textbf{hierarchical cascade}: (1)~after physics simulation, a \texttt{set\_state\_completion} signal propagates top-down; (2)~each agent requests its updated local state from the ProxyAgent; (3)~each agent computes \texttt{compute\_local\_reward(local\_state)} from its visibility-filtered view; (4)~parent agents apply \texttt{reward\_delay} before computing their own rewards, ensuring subordinate rewards are available for hierarchical credit assignment. This reflects real CPS reward semantics: a grid operator's cost aggregates from device-level metrics with natural propagation delays.

%==============================================================================
\section{\heron{} Framework}
\label{sec:framework}
%==============================================================================

\heron{} is organized around six modules: \textbf{Agents} (hierarchical decision-makers), \textbf{ProxyAgent} (state mediation), \textbf{Features} (visibility-labeled observations), \textbf{Messaging} (channel-based communication), \textbf{Protocols} (coordination mechanisms), and \textbf{Environments} (physics backends).

\subsection{Architecture Overview}
\label{sec:overview}

Figure~\ref{fig:architecture} shows the architecture. The key invariant is that \textbf{no agent directly accesses environment state}. All state reads and writes are mediated by the ProxyAgent, which enforces visibility constraints.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    node distance=0.6cm,
    box/.style={rectangle, draw, minimum width=2.0cm, minimum height=0.7cm, align=center, font=\scriptsize},
    arrow/.style={->, >=stealth, thick},
    dasharrow/.style={->, >=stealth, thick, dashed}
]
% RL Layer
\node[box, fill=blue!15] (rl) {RL Training\\(RLlib/SB3)};

% Environment Layer
\node[box, fill=green!15, below=of rl] (env) {\heron{} Env\\(\texttt{step} / \texttt{run\_event\_driven})};

% ProxyAgent
\node[box, fill=red!15, below left=0.7cm and 1.2cm of env] (proxy) {ProxyAgent\\(State + Visibility)};

% Message Broker
\node[box, fill=yellow!25, below right=0.7cm and 1.2cm of env] (broker) {MessageBroker\\(Channels)};

% System Agent
\node[box, fill=cyan!15, below=1.8cm of env] (sys) {SystemAgent (L3)};

% Coordinator Agents
\node[box, fill=orange!20, below left=0.6cm and 0.7cm of sys] (ca1) {Coordinator\\Agent (L2)};
\node[box, fill=orange!20, below right=0.6cm and 0.7cm of sys] (ca2) {Coordinator\\Agent (L2)};

% Field Agents
\node[box, fill=purple!15, below=0.5cm of ca1, minimum width=1.6cm] (f1) {FieldAgents (L1)};
\node[box, fill=purple!15, below=0.5cm of ca2, minimum width=1.6cm] (f2) {FieldAgents (L1)};

% Physics
\node[box, fill=gray!20, below=3.2cm of env, xshift=-3.5cm] (physics) {Physics Backend};

% EventScheduler
\node[box, fill=teal!20, right=0.3cm of broker] (sched) {Event\\Scheduler};

% Arrows
\draw[arrow] (rl) -- (env);
\draw[arrow] (env) -- (proxy);
\draw[arrow] (env) -- (broker);
\draw[dasharrow] (proxy) -- node[left, font=\tiny] {filtered obs} (sys);
\draw[dasharrow] (broker) -- node[right, font=\tiny] {msgs} (sys);
\draw[arrow] (sys) -- (ca1);
\draw[arrow] (sys) -- (ca2);
\draw[arrow] (ca1) -- (f1);
\draw[arrow] (ca2) -- (f2);
\draw[dasharrow] (f1.south) |- (physics.east);
\draw[dasharrow] (f2.south) |- (physics.east);
\draw[dasharrow] (sched) -- node[above, font=\tiny] {ticks} (broker);

% Protocol annotation
\node[above=0.1cm of broker, font=\tiny, text=gray] (proto) {Protocols};

\end{tikzpicture}
\caption{\heron{} architecture. The ProxyAgent mediates all state access with visibility filtering. Agents are organized in a three-level hierarchy with composable protocols. The EventScheduler drives event-driven execution.}
\label{fig:architecture}
\end{figure}

In the environment-centric paradigm (PettingZoo, RLlib), the environment class owns observation construction, action routing, reward logic, coordination, \emph{and} physics. In \heron{}'s \emph{agent-centric} paradigm, the environment owns only the physics bridge; agents carry their own features, visibility declarations, reward functions, and timing. This separation ensures that coordination changes do not become physics changes.

%------------------------------------------------------------------------------
\subsection{Three-Level Agent Hierarchy}
\label{sec:hierarchy}
%------------------------------------------------------------------------------

Many CPS are naturally hierarchical. \heron{} formalizes this:
\begin{itemize}[nosep]
    \item \textbf{Level 1 (FieldAgent)}: Local sensing and actuation. Access to \texttt{owner}+\texttt{public} features. Examples: individual devices, sensors, traffic signals. Default tick: 1\,s.
    \item \textbf{Level 2 (CoordinatorAgent)}: Manages subordinate FieldAgents via protocols. Access to \texttt{upper\_level}+\texttt{owner}+\texttt{public}. Examples: zone controllers, microgrid aggregators. Default tick: 60\,s.
    \item \textbf{Level 3 (SystemAgent)}: Global coordination, drives simulation. Access to \texttt{system} (all scopes). Examples: central operators, network controllers. Default tick: 300\,s.
\end{itemize}

Each agent carries: (1)~\textbf{State} composed of \texttt{FeatureProvider} instances with declared visibility; (2)~\textbf{Action} with mixed continuous/discrete specification; (3)~\textbf{Reward} via \texttt{compute\_local\_reward(local\_state)}; (4)~\textbf{Timing} via \texttt{TickConfig} (tick interval, delays, jitter).

\paragraph{Dual execution modes.}
\heron{} supports two execution patterns over the same agent hierarchy:
\begin{enumerate}[nosep]
    \item \textbf{Synchronous} (\texttt{env.step(actions)}): Direct \texttt{execute()} calls; agents sync state from ProxyAgent, then act. CTDE-compatible.
    \item \textbf{Event-driven} (\texttt{env.run\_event\_driven(t\_end)}): Discrete-event simulation via heap-based \texttt{EventScheduler} with seven event types (\texttt{AGENT\_TICK}, \texttt{ACTION\_EFFECT}, \texttt{MESSAGE\_DELIVERY}, \texttt{OBSERVATION\_READY}, \texttt{ENV\_UPDATE}, \texttt{SIMULATION}, \texttt{CUSTOM}). Agents tick independently at heterogeneous rates with configurable delays and jitter.
\end{enumerate}

Algorithm~\ref{alg:event-driven} outlines the event-driven execution loop.

\begin{algorithm}[t]
\caption{Event-driven execution via \texttt{EventScheduler}}
\label{alg:event-driven}
\begin{algorithmic}[1]
\REQUIRE EventScheduler $\mathcal{S}$ (heap-based priority queue), ProxyAgent $\mathcal{P}$
\REQUIRE Agents with TickConfig (tick\_interval, obs\_delay, act\_delay, msg\_delay)
\STATE
\STATE \textbf{function} \textsc{RunEventDriven}($t_\text{end}$, event\_analyzer):
\STATE \hspace{0.5cm} \textbf{for each} event $\in$ $\mathcal{S}$.run\_until($t_\text{end}$):
\STATE \hspace{1.0cm} \textbf{match} event.type:
\STATE
\STATE \hspace{1.0cm} \textbf{case} AGENT\_TICK:
\STATE \hspace{1.5cm} agent.tick($\mathcal{S}$, event.timestamp) \COMMENT{Set time, check upstream}
\STATE \hspace{1.5cm} obs $\gets$ $\mathcal{P}$.get\_observation(agent\_id, protocol) \COMMENT{Visibility-filtered}
\STATE \hspace{1.5cm} action $\gets$ agent.compute\_action(obs) \COMMENT{Upstream $>$ policy}
\STATE \hspace{1.5cm} $\mathcal{S}$.schedule(ACTION\_EFFECT, $t$ + act\_delay)
\STATE \hspace{1.5cm} $\mathcal{S}$.schedule(AGENT\_TICK, $t$ + tick\_interval) \COMMENT{Next tick}
\STATE
\STATE \hspace{1.0cm} \textbf{case} ACTION\_EFFECT:
\STATE \hspace{1.5cm} agent.apply\_action() \COMMENT{Physical effect}
\STATE \hspace{1.5cm} $\mathcal{P}$.set\_local\_state(agent\_id, new\_state)
\STATE
\STATE \hspace{1.0cm} \textbf{case} MESSAGE\_DELIVERY:
\STATE \hspace{1.5cm} broker.deliver(sender, recipient, payload)
\STATE
\STATE \hspace{1.0cm} event\_analyzer.analyze(event) \COMMENT{User-defined hook}
\end{algorithmic}
\end{algorithm}

%------------------------------------------------------------------------------
\subsection{ProxyAgent: Centralized State Mediation}
\label{sec:proxy}
%------------------------------------------------------------------------------

The \textbf{ProxyAgent} is a non-learning singleton that stores all agent states and mediates every observation request through visibility filtering:
\begin{itemize}[nosep]
    \item \textbf{State storage}: All \texttt{State} objects (containing \texttt{FeatureProvider} instances) are stored in the ProxyAgent, not in agents themselves.
    \item \textbf{Visibility enforcement}: \texttt{get\_observation(sender\_id, protocol)} filters features by visibility labels, requester identity, and hierarchy level.
    \item \textbf{Global aggregation}: \texttt{get\_global\_states(sender\_id)} returns visibility-filtered views of all agents, enabling coordinators to see subordinate aggregates without private fields.
\end{itemize}

The ProxyAgent enforces a key invariant: \emph{an agent can only observe features whose visibility labels include a scope accessible to that agent's hierarchy level}. This prevents the ``global state leak'' common in MARL benchmarks.

\begin{table}[t]
\centering
\caption{Visibility scopes in \heron{}.}
\label{tab:visibility}
\small
\begin{tabular}{@{}lp{3.0cm}p{6.6cm}@{}}
\toprule
\textbf{Scope} & \textbf{Who can access} & \textbf{Example signals} \\
\midrule
\texttt{public} & All agents & System time, global alerts, cost/safety metrics \\
\texttt{owner} & Owning agent only & Device SOC, internal costs, local setpoints \\
\texttt{upper\_level} & Parent in hierarchy & Subordinate aggregates, boundary bus voltages \\
\texttt{system} & System-level (L3) agents & Full network state, aggregate gen/load, inter-area flows \\
\bottomrule
\end{tabular}
\end{table}

This enables \textbf{systematic observability ablations}: progressively restrict visibility (\texttt{system} $\rightarrow$ \texttt{upper\_level} $\rightarrow$ \texttt{owner}) to quantify minimum information requirements---by changing a single \texttt{visibility} class attribute on each \texttt{FeatureProvider}, with no environment code changes.

%------------------------------------------------------------------------------
\subsection{Composable Observability with FeatureProviders}
\label{sec:observability}
%------------------------------------------------------------------------------

Observations are assembled from composable \textbf{FeatureProviders}---dataclasses that map a slice of state into a feature vector and declare visibility scopes via the \texttt{FeatureMeta} metaclass. Example:

\begin{lstlisting}
class BatterySOC(FeatureProvider):
    visibility: ClassVar = ["owner"]      # Only owning agent
    soc: float = 0.5
    capacity_kwh: float = 100.0

class SystemFrequency(FeatureProvider):
    visibility: ClassVar = ["system"]     # Only SystemAgent
    frequency_hz: float = 50.0
\end{lstlisting}

To change what agents observe, modify the \texttt{visibility} attribute---no environment code changes needed.

%------------------------------------------------------------------------------
\subsection{MessageBroker for Explicit Communication}
\label{sec:messaging}
%------------------------------------------------------------------------------

\heron{} decouples agent communication from observation through a \textbf{MessageBroker} with six message types (\texttt{ACTION}, \texttt{INFO}, \texttt{BROADCAST}, \texttt{STATE\_UPDATE}, \texttt{RESULT}, \texttt{CUSTOM}). Channels are named with convention \texttt{env\_\{id\}\_\_action\_\_\{sender\}\_to\_\{recipient\}}, providing environment isolation for multi-environment parallel training. The default \texttt{InMemoryBroker} supports single-process training; the interface is extensible to distributed backends.

%------------------------------------------------------------------------------
\subsection{Protocol System for Coordination}
\label{sec:protocols}
%------------------------------------------------------------------------------

\heron{} separates coordination semantics from physics through a modular protocol system. Each \texttt{Protocol} composes:
\begin{enumerate}[nosep]
    \item \textbf{CommunicationProtocol}: What messages are exchanged (\eg state sharing, price signals).
    \item \textbf{ActionProtocol}: How a coordinator's action decomposes into subordinate actions.
\end{enumerate}

Built-in families: \textbf{VerticalProtocol} (hierarchical: \texttt{NoCommunication} + \texttt{VectorDecomposition}---splits joint action into per-subordinate segments), \textbf{HorizontalProtocol} (peer-to-peer: \texttt{StateShare} + \texttt{NoAction}). Domain-specific protocols (\eg setpoint dispatch, market clearing, consensus) are subclasses. Under identical observability, researchers compare coordination strategies by swapping protocols at configuration time.

%==============================================================================
\section{Event-Driven Execution: The Five Mismatches}
\label{sec:mismatches}
%==============================================================================

Classical MARL assumes policies operate under well-behaved MDP dynamics. In event-driven execution with hierarchical control, five categories of mismatch arise---each violating specific MDP assumptions.

\paragraph{1. Stale observations (timing-induced).}
Agents at different hierarchy levels tick at different rates. A coordinator observing at 1\,Hz cannot see the 100 micro-actions a field agent has taken since its last tick. The coordinator's state transition model assumes a static world between decisions---violated by continuous subordinate activity. \emph{In \heron{}}: configurable via per-agent \texttt{TickConfig.tick\_interval}.

\paragraph{2. Partial observability (asymmetric information).}
Each agent sees only features matching its visibility scope. Agent A detects a local event invisible to Agent B, leading to conflicting decisions. \emph{In \heron{}}: modeled via \texttt{FeatureProvider.visibility}---switching partial to full observability is a one-line class attribute change.

\paragraph{3. Communication failure (message loss).}
A directive from a coordinator to a subordinate may be delayed or dropped. Agents operate on incompatible plans. \emph{In \heron{}}: \texttt{MessageBroker} with configurable \texttt{msg\_delay} and optional loss probability.

\paragraph{4. State corruption (external disruptions).}
An external event disrupts an agent's state, but the coordinator's next observation reflects pre-disruption conditions. \emph{In \heron{}}: the ProxyAgent's state cache naturally models stale entries.

\paragraph{5. Goal conflict (reward misalignment).}
Hierarchical agents may optimize different objectives (speed vs.\ safety). Actions are locally rational but globally incompatible. \emph{In \heron{}}: each agent defines its own \texttt{compute\_local\_reward()}.

\begin{table}[t]
\centering
\caption{How \heron{} makes each mismatch a systematic experimental variable.}
\label{tab:mismatches}
\small
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Mismatch} & \textbf{Control Mechanism} & \textbf{Configuration} \\
\midrule
Stale observations & \texttt{TickConfig.tick\_interval} & Different rates per level \\
Partial observability & \texttt{FeatureProvider.visibility} & \texttt{public/owner/upper/system} \\
Communication failure & \texttt{MessageBroker} + \texttt{msg\_delay} & Delay distributions \\
State corruption & ProxyAgent state cache & Custom event injection \\
Goal conflict & Per-agent \texttt{compute\_local\_reward()} & Different reward functions \\
\bottomrule
\end{tabular}
\end{table}

The experimental protocol: train with \texttt{env.step()} (all mismatches absent), evaluate with \texttt{env.run\_event\_driven()} (mismatches present at configured levels), and quantify the gap.

%==============================================================================
\section{Related Work}
\label{sec:related}
%==============================================================================

\textbf{Multi-agent benchmarks.}
MPE~\cite{mordatch2018emergence}, SMAC~\cite{samvelyan2019starcraft}, and SMACv2 inherit synchronous stepping with globally derived observations. PettingZoo~\cite{terry2021pettingzoo} provides important API standardization (AEC model, reward-attribution fixes) but its abstractions do not model hierarchical information flow, event-driven execution, or coordination protocols. EPyMARL~\cite{papoudakis2021benchmarking} and MARLlib~\cite{hu2023marllib} focus on algorithm benchmarking---orthogonal to \heron{}'s focus on environment construction.

\textbf{CPS-domain RL benchmarks.}
Grid2Op~\cite{donnot2020introducing} and gym-anm~\cite{henry2021gym} are single-agent. CityLearn~\cite{vazquez2019citylearn} and PowerGridworld~\cite{biagioni2022powergridworld} support multi-agent settings but offer only binary observability and no protocol abstraction. EnEnv (AAMAS 2025) provides a rigorous MARL benchmark for battery storage with real-world physics, but inherits PettingZoo's synchronous model---no event-driven mode, implicit coordination via shared rewards.

\textbf{Discrete-event multi-agent simulation.}
ABIDES~\cite{byrd2020abides} pioneers event-driven multi-agent simulation for financial markets. However, ABIDES is event-driven \emph{only} (no synchronous CTDE mode), uses a flat agent model (no hierarchy), and provides no visibility framework or protocol system. ABIDES-MARL (2025) extends to multiple RL agents but re-introduces synchronization barriers, partially negating asynchronous benefits. \heron{} provides dual-mode execution, hierarchical agents, and systematic information control---none present in ABIDES.

\textbf{Hierarchical RL and communication.}
Feudal networks~\cite{vezhnevets2017feudal}, DIAL/CommNet~\cite{foerster2016learning}, and TarMAC~\cite{das2019tarmac} study \emph{learning} hierarchy or communication. \heron{} provides a substrate where hierarchy and communication topology are \emph{configurable experimental variables}.

\begin{table}[t]
\centering
\caption{Comparison with related frameworks. \heron{} uniquely combines dual-mode execution, hierarchy, multi-level visibility, composable protocols, and centralized state mediation.}
\label{tab:comparison}
\small
\begin{tabular}{@{}lcccccc@{}}
\toprule
\textbf{Framework} & \textbf{Event} & \textbf{Dual} & \textbf{Hierarchy} & \textbf{Visibility} & \textbf{Protocols} & \textbf{State Med.} \\
\midrule
PettingZoo & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark \\
EPyMARL & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark \\
SMAC & \xmark & \xmark & \xmark & Partial & \xmark & \xmark \\
Grid2Op & \xmark & \xmark & \xmark & N/A & \xmark & \xmark \\
CityLearn & \xmark & \xmark & \xmark & Binary & \xmark & \xmark \\
ABIDES & \cmark & \xmark & \xmark & Implicit & \xmark & \xmark \\
\midrule
\textbf{\heron{}} & \cmark & \cmark & \cmark & \textbf{4-level} & \cmark & \cmark \\
\bottomrule
\end{tabular}
\end{table}

\heron{} complements PettingZoo (different abstraction level---what happens \emph{inside} the environment rather than the env-algorithm interface) and EPyMARL/MARLlib (environment construction rather than algorithm benchmarking). \heron{} environments can export to PettingZoo's API for algorithm compatibility.

%==============================================================================
\section{API and Usage}
\label{sec:api}
%==============================================================================

\subsection{Entry Points}

\heron{} exposes two primary execution APIs:

\paragraph{Synchronous training.}
\begin{lstlisting}
obs, infos = env.reset(seed=42)
for step in range(max_steps):
    actions = {aid: policy(obs[aid]) for aid in obs}
    obs, rewards, terminated, truncated, infos = env.step(actions)
    if terminated["__all__"] or truncated["__all__"]:
        break
\end{lstlisting}

\texttt{env.step()} delegates to \texttt{SystemAgent.execute()}, which processes actions top-down through the hierarchy, runs physics, collects visibility-filtered observations, and computes per-agent rewards.

\paragraph{Event-driven evaluation.}
\begin{lstlisting}
# Configure realistic timing
for agent in env.registered_agents.values():
    if isinstance(agent, FieldAgent):
        agent.tick_config = TickConfig.with_jitter(
            tick_interval=1.0, obs_delay=0.1, act_delay=0.2,
            jitter_type=JitterType.GAUSSIAN, jitter_ratio=0.2, seed=42)

obs, infos = env.reset(seed=42, jitter_seed=42)
result = env.run_event_driven(
    event_analyzer=EventAnalyzer(), t_end=3600.0)
\end{lstlisting}

\subsection{Building Environments}

\heron{} provides a fluent \texttt{EnvBuilder}:
\begin{lstlisting}
env = (EnvBuilder("my_env")
    .add_agent("dev_0", MyDevice, features=[BatterySOC()])
    .add_agent("dev_1", MyDevice, features=[BatterySOC()])
    .add_coordinator("zone_0",
        subordinates=["dev_0", "dev_1"],
        protocol=VerticalProtocol())
    .add_system_agent(tick_config=TickConfig(tick_interval=300.0))
    .simulation(my_physics_fn)
    .build())
\end{lstlisting}

The builder auto-creates the hierarchy, resolves subordinates, and wires up ProxyAgent, MessageBroker, and EventScheduler.

\subsection{RLlib Integration}

\texttt{RLlibBasedHeronEnv} wraps any \heron{} environment as an RLlib \texttt{MultiAgentEnv}:
\begin{lstlisting}
config = (PPOConfig()
    .environment(env=RLlibBasedHeronEnv, env_config={
        "agents": [{"agent_id": "bat_0", "agent_cls": BatteryAgent,
                     "features": [BatterySOC()],
                     "coordinator": "mg_0"}],
        "coordinators": [{"coordinator_id": "mg_0",
                          "protocol": VerticalProtocol()}],
        "simulation": pandapower_sim, "max_steps": 100}))
algo = config.build()
algo.train()
\end{lstlisting}

The adapter handles converting \heron{} \texttt{Observation} objects to flat numpy vectors, exposing per-agent spaces, and episode truncation. For event-driven evaluation during training, use \texttt{HeronEnvRunner}:
\begin{lstlisting}
config = config.env_runners(
    env_runner_cls=HeronEnvRunner,
    evaluation_config=HeronEnvRunner.evaluation_config(t_end=3600.0))
\end{lstlisting}

%==============================================================================
\section{Power Systems Case Study}
\label{sec:power}
%==============================================================================

We instantiate \heron{} for multi-microgrid coordination using PandaPower AC power flow, demonstrating a concrete CPS benchmark where information structure and coordination protocols matter.

\paragraph{Provided assets.}
\begin{itemize}[nosep]
    \item \textbf{Standard networks}: IEEE 13/34/123-bus, CIGRE LV, Case34 3-phase, Case LVMG.
    \item \textbf{Device models}: Generators, ESS, tap-changing transformers, inverter-based sources, shunt capacitors.
    \item \textbf{FeatureProviders}: 20 domain-specific providers (Table~\ref{tab:features}).
    \item \textbf{Tutorials}: 7 notebooks covering single microgrid, multi-microgrid, event-driven, and CTDE training.
\end{itemize}

\begin{table}[t]
\centering
\caption{FeatureProviders in the power systems case study (20 total). ``Vis.'' indicates implemented visibility; others default to unrestricted.}
\label{tab:features}
\small
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Category} & \textbf{FeatureProvider} & \textbf{Vis.} & \textbf{Description} \\
\midrule
\multirow{6}{*}{\rotatebox{90}{\scriptsize Device}} & ElectricalBasePh & -- & Active/reactive power, voltage \\
& PowerLimits & -- & Min/max generation capacity \\
& StorageBlock & owner & SOC, charge/discharge limits \\
& StatusBlock & -- & Device availability \\
& TapChangerPh & -- & Tap position and limits \\
& InverterBasedSource & -- & Inverter mode, reactive cap. \\
\midrule
\multirow{4}{*}{\rotatebox{90}{\scriptsize Net.}} & BusVoltages & -- & Voltage magnitudes/angles \\
& LineFlows & -- & Branch flows, loading \% \\
& ThermalLoading & -- & Line thermal limits \\
& PhaseConnection & -- & Bus connection info \\
\midrule
\multirow{5}{*}{\rotatebox{90}{\scriptsize System}} & SystemFrequency & sys, upper & System-wide frequency \\
& AggregateGeneration & sys & Total generation \\
& AggregateLoad & sys & Total load \\
& InterAreaFlows & sys & Inter-area power flows \\
& SystemImbalance & sys, upper & Gen-load imbalance \\
\midrule
\multirow{3}{*}{\rotatebox{90}{\scriptsize Ctrl}} & VoltVarCurve & -- & Volt-VAR curves \\
& ShuntCapacitorBlock & -- & Capacitor bank state \\
& NetworkMetrics & -- & Total gen, load, losses \\
\midrule
\multirow{2}{*}{\rotatebox{90}{\scriptsize Meta}} & StepState & -- & Timestep, progress \\
& CostSafetyMetrics & public & Cost and safety metrics \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Agent hierarchy.}
\textbf{DeviceAgent} (FieldAgent): individual DERs---\texttt{Generator}, \texttt{ESS}, \texttt{Transformer} subclasses. \textbf{PowerGridAgent} (CoordinatorAgent): microgrid management, power flow integration. \textbf{GridSystemAgent} (SystemAgent): network-wide coordination.

\paragraph{Benchmark questions enabled.}
\begin{itemize}[nosep]
    \item \textbf{Minimum observability}: What visibility scope is sufficient for stable voltage regulation?
    \item \textbf{Information asymmetry}: Can microgrids coordinate with only upper-level aggregates?
    \item \textbf{Protocol sensitivity}: Under identical visibility, which coordination mechanism is more robust?
    \item \textbf{Sim-to-real gap}: Do synchronously trained policies degrade under event-driven execution with SCADA-calibrated timing?
\end{itemize}

%==============================================================================
\section{Experiments}
\label{sec:experiments}
%==============================================================================

We present experiments demonstrating \heron{}'s benchmark capabilities. Values marked with $\dagger$ are preliminary.

\subsection{Observability Ablation}

\begin{table}[h]
\centering
\caption{Observability ablation (IEEE 34-bus, 3 microgrids).}
\label{tab:observability-ablation}
\small
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Observability} & \textbf{Cost (\$)} & \textbf{Degrad.} & \textbf{Safety (\%)} & \textbf{Conv.\ (ep)} \\
\midrule
System (full) & \exampleval{859} & Baseline & \exampleval{2.1} & \exampleval{2400} \\
Upper-level & \exampleval{891} & \exampleval{+3.7\%} & \exampleval{2.8} & \exampleval{2600} \\
Owner-only & \exampleval{1024} & \exampleval{+19\%} & \exampleval{8.7} & \exampleval{4200} \\
Public-only & \exampleval{1543} & \exampleval{+80\%} & \exampleval{24} & Fails \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Sim-to-Real Gap (Train/Test Mismatch)}

\begin{table}[h]
\centering
\caption{Synchronous-to-event-driven performance gap.}
\label{tab:mismatch}
\small
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Train Vis.} & \textbf{Test Vis.} & \textbf{Cost (\$)} & \textbf{Degrad.} \\
\midrule
System & System & \exampleval{859} & -- \\
Upper-level & Upper-level & \exampleval{891} & +3.7\% \\
\midrule
\textbf{System} & \textbf{Upper-level} & \exampleval{1055} & \textbf{+23\%} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Protocol Comparison}

\begin{table}[h]
\centering
\caption{Protocol comparison (upper-level visibility, 3 microgrids).}
\label{tab:protocol}
\small
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Protocol} & \textbf{Cost (\$)} & \textbf{Overhead} & \textbf{Info Shared} \\
\midrule
Vertical (setpoint) & \exampleval{891} & Low & Full control \\
Vertical (price signal) & \exampleval{912} & Medium & Prices only \\
Horizontal (state share) & \exampleval{928} & High & Neighbor states \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Event-Driven Timing Sensitivity}

\begin{table}[h]
\centering
\caption{Timing sensitivity (trained synchronously, evaluated with delays).}
\label{tab:timing}
\small
\begin{tabular}{@{}llcc@{}}
\toprule
\textbf{Config} & \textbf{Distribution} & \textbf{Degrad.} & \textbf{Safety} \\
\midrule
Synchronous & -- & Baseline & \exampleval{2.1\%} \\
Uniform delay & $U(0, 2\text{s})$ & \exampleval{+8\%} & \exampleval{4.3\%} \\
SCADA-calibrated & LogNormal($\mu$=2s, $\sigma$=0.8s) & \exampleval{+15\%} & \exampleval{6.1\%} \\
Gaussian jitter & $\mathcal{N}(0, 0.5\text{s})$ + base & \exampleval{+5\%} & \exampleval{3.2\%} \\
Heterogeneous & Per-agent different $\tau$ & \exampleval{+22\%} & \exampleval{8.5\%} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Algorithm Comparison}

\begin{table}[h]
\centering
\caption{Algorithm comparison under system vs.\ owner-only visibility.}
\label{tab:algorithms}
\small
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Algorithm} & \textbf{Sys Cost} & \textbf{Owner Cost} & \textbf{Degrad.} & \textbf{Category} \\
\midrule
MAPPO & \exampleval{859} & \exampleval{1024} & \exampleval{+19\%} & Policy gradient \\
IPPO & \exampleval{903} & \exampleval{1089} & \exampleval{+21\%} & Independent \\
QMIX & \todo{TBD} & \todo{TBD} & \todo{TBD} & Value decomp. \\
TarMAC & \todo{TBD} & \todo{TBD} & \todo{TBD} & Communication \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Hypothesis}: The relative ordering (system $>$ upper-level $>$ owner) holds across algorithm families, validating that observability effects are algorithm-agnostic properties of the benchmark.

%==============================================================================
\section{Conclusion}
\label{sec:conclusion}
%==============================================================================

We presented \heron{}, a hierarchical multi-agent framework built on an agent-centric design principle: \emph{agents own the coordination; the environment owns only the physics}. This separation makes execution model, information structure, and coordination mechanisms explicit as benchmark variables---without modifying the physics backend. Through a ProxyAgent mediating all state access, FeatureProviders with four-level visibility, a channel-based MessageBroker, and composable protocols, \heron{} enables systematic studies of how information constraints and timing realism affect MARL in cyber-physical systems. Dual-mode execution (synchronous CTDE for training, event-driven for deployment validation) provides a direct measurement of the simulation-to-deployment gap. A power systems case study demonstrates these capabilities concretely. \heron{} is released open-source to support reproducible research on event-driven, information-constrained multi-agent coordination.

%==============================================================================
% References
%==============================================================================

\bibliographystyle{plainnat}
\bibliography{references}

%==============================================================================
% Appendix
%==============================================================================
\newpage
\appendix

\section{Core API Reference}
\label{app:api}

\begin{table}[h]
\centering
\caption{Core \heron{} API summary.}
\label{tab:api}
\small
\begin{tabular}{@{}lp{8.5cm}@{}}
\toprule
\textbf{Method} & \textbf{Signature and Description} \\
\midrule
\multicolumn{2}{l}{\textit{Environment (EnvCore)}} \\
\texttt{step} & \texttt{step(actions: Dict) -> (obs, rewards, term, trunc, info)} \\
\texttt{run\_event\_driven} & \texttt{run\_event\_driven(analyzer, t\_end) -> EpisodeResult} \\
\texttt{reset} & \texttt{reset(seed, jitter\_seed) -> (obs, infos)} \\
\midrule
\multicolumn{2}{l}{\textit{Agent}} \\
\texttt{tick} & \texttt{tick(scheduler, current\_time) -> None} \\
\texttt{execute} & \texttt{execute(actions, proxy) -> None} \\
\texttt{compute\_action} & \texttt{compute\_action(obs, scheduler) -> action} \\
\texttt{apply\_action} & \texttt{apply\_action() -> None} \\
\texttt{compute\_local\_reward} & \texttt{compute\_local\_reward(local\_state) -> float} \\
\midrule
\multicolumn{2}{l}{\textit{ProxyAgent}} \\
\texttt{get\_observation} & \texttt{get\_observation(sender\_id, protocol) -> Observation} \\
\texttt{get\_local\_state} & \texttt{get\_local\_state(sender\_id, protocol) -> Dict} \\
\texttt{get\_global\_states} & \texttt{get\_global\_states(sender\_id, protocol) -> Dict} \\
\texttt{set\_local\_state} & \texttt{set\_local\_state(agent\_id, state) -> None} \\
\midrule
\multicolumn{2}{l}{\textit{EventScheduler}} \\
\texttt{run\_until} & \texttt{run\_until(t\_end, max\_events) -> Iterable[Event]} \\
\texttt{schedule\_agent\_tick} & \texttt{schedule\_agent\_tick(agent\_id, timestamp)} \\
\texttt{schedule\_action\_effect} & \texttt{schedule\_action\_effect(agent\_id, delay)} \\
\texttt{schedule\_message\_delivery} & \texttt{schedule\_message\_delivery(sender, recip, msg, delay)} \\
\midrule
\multicolumn{2}{l}{\textit{Protocol}} \\
\texttt{coordinate} & \texttt{coordinate(state, action, context) -> (msgs, actions)} \\
\bottomrule
\end{tabular}
\end{table}

\section{TickConfig Defaults}
\label{app:tickconfig}

\begin{table}[h]
\centering
\caption{Default timing configurations per agent level.}
\small
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Agent Level} & \textbf{Default Tick} & \textbf{Jitter Types} & \textbf{Description} \\
\midrule
FieldAgent (L1) & 1.0\,s & \textsc{None}, Uniform, Gaussian & Fast local control \\
CoordinatorAgent (L2) & 60.0\,s & \textsc{None}, Uniform, Gaussian & Zone management \\
SystemAgent (L3) & 300.0\,s & \textsc{None}, Uniform, Gaussian & Supervisory review \\
\bottomrule
\end{tabular}
\end{table}

Each \texttt{TickConfig} parameter (\texttt{tick\_interval}, \texttt{obs\_delay}, \texttt{act\_delay}, \texttt{msg\_delay}, \texttt{reward\_delay}) supports independent jitter. \texttt{UNIFORM}: $\text{base} \pm \text{base} \times \text{ratio}$. \texttt{GAUSSIAN}: $\mathcal{N}(\text{base}, \text{base} \times \text{ratio})$. All values are clamped to a non-negative floor.

\section{Network Specifications}
\label{app:networks}

\begin{table}[h]
\centering
\caption{Test network specifications (power case study).}
\small
\begin{tabular}{@{}lccccl@{}}
\toprule
\textbf{Network} & \textbf{Buses} & \textbf{Lines} & \textbf{Voltage (kV)} & \textbf{Peak (MW)} & \textbf{Phases} \\
\midrule
IEEE 13-Bus & 13 & 11 & 4.16 & 3.5 & 3-phase \\
IEEE 34-Bus & 34 & 33 & 24.9 & 1.77 & 3-phase \\
IEEE 123-Bus & 123 & 127 & 4.16 & 3.49 & 3-phase \\
CIGRE LV & 14 & 13 & 0.4 & 0.245 & 3-phase \\
Case34 3ph & 34 & 33 & 24.9 & 1.77 & Unbalanced \\
Case LVMG & \todo{TBD} & \todo{TBD} & 0.4 & \todo{TBD} & 3-phase \\
\bottomrule
\end{tabular}
\end{table}

\section{Hyperparameters}
\label{app:hyperparams}

\begin{table}[h]
\centering
\caption{MAPPO training hyperparameters.}
\small
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
Learning rate & $3 \times 10^{-4}$ \\
Discount ($\gamma$) & 0.99 \\
GAE $\lambda$ & 0.95 \\
Clip $\epsilon$ & 0.2 \\
Entropy coeff. & 0.01 \\
Value loss coeff. & 0.5 \\
Max grad norm & 0.5 \\
Batch / minibatch & 4096 / 256 \\
PPO epochs & 10 \\
Workers & 4 \\
Network & [256, 256] MLP, ReLU \\
\bottomrule
\end{tabular}
\end{table}

\section{Safety Metrics}
\label{app:safety}

\begin{align}
V_{\text{voltage}} &= \sum_{b \in \mathcal{B}} \left[\max(0, V_b - 1.05) + \max(0, 0.95 - V_b)\right] \\
V_{\text{loading}} &= \sum_{l \in \mathcal{L}} \max(0, \text{Loading}_l - 1.0) \\
V_{\text{SOC}} &= \sum_{s \in \mathcal{S}} \left[\max(0, 0.1 - \text{SOC}_s) + \max(0, \text{SOC}_s - 0.9)\right]
\end{align}

Voltage limits: 0.95--1.05 p.u. Line loading limit: 100\%. SOC safe range: 10--90\%.

\section{Code and Reproducibility}
\label{app:code}

\heron{} is available at: \url{https://github.com/Criss-Wang/PowerGym}

\begin{verbatim}
pip install -e ".[all]"       # Full installation
pip install -e ".[powergrid]" # Power grid case study only
\end{verbatim}

\todo{Add Croissant metadata record for NeurIPS D\&B compliance.}

\end{document}
